---
title: "Brownlows Bets"
author: "Shaun I"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings
## the standard deviation and a 95% confidence interval for their earnings?
  
  
```{r}
results = matrix(rep(0, 500))
for(i in 1:500){
  results[i] = BrLw(35, 2, 637)
  
}
hist(results)
print(paste("Mean profits = ", round(mean(results))))
#95 percent confidence interval
sortedResults = sort(results)

lower_bound <- quantile(results, 0.025)
upper_bound <- quantile(results, 0.975)
print("95% confidence Interval")
print(c(lower_bound, upper_bound))



```
Gets the 2.5th and 97.5 percentiles
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->


# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 

```{r}
optimize = matrix(rep(0, 100))

for (i in seq(0, 10, by = 0.1)) {
  

  
  # Update the optimize matrix
  optimize[round(i*10)] <- BrLw(35, i, 637)


  
}
min_value <- min(abs(optimize))
min_index <- which(abs(optimize) == min_value, arr.ind = TRUE)

print(paste("Closest value to zero:", min_value))
print(paste("Index position:", min_index))

```
If Brownlow wagers 4.6 dollars over the student's 1 dollar, then the odds would be roughly even with the closest value to zero given 637 epochs being 4.2 dollars.
<!--Decide on your methods:  In this section explain the approach you will take to attack the issue  --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explainhow you construct a 95% confidence interval about your estimate of his winnings !-->
```{r}
optimize
```




# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
optimize = matrix(rep(3, 100))

for (i in 1:100) {

  optimize[i] = BrLw(i, 2, 637)
  
}
min_value <- min(abs(optimize))
min_index <- which(abs(optimize) == min_value, arr.ind = TRUE)

print(paste("Closest value to zero:", min_value))
print(paste("Index position:", min_index))

```

The number of students that tunes Brownlow's earnings the closest to 0 is 28 students with a 2:1 bet with a mean earning of 7$.
```{r}
optimize
```




# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

If Dr. Brownlow was consistently losing money he can either increase the size of his class or wager less money. If Dr. Brownlow was consistently wining money he can either decrease the size of his class or wager more money to have a net zero profit. Based on the simulations done above, I found that the number of students that tunes Brownlow's earnings the closest to 0 is 28 students with a 2:1 bet with a mean earning of 7$. Furthermore, I found that if Brownlow wagers 4.6 dollars over the student's 1 dollar, then the odds would be roughly even with the closest value to zero given 637 epochs being 4.2 dollars.